\documentclass{article}
\usepackage[hang,bf,small]{caption}
\usepackage{hyperref}
\usepackage[all]{hypcap}

\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.125in}
\setlength{\textwidth}{6.25in}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{amssymb}


\begin{document}
\begin{flushleft}
Jose Garmilla \\
David Vartanyan\\
Thermodynamics of Circumbinary Disks\\
\today\\
\end{flushleft}

\section{Introduction}
We consider the evolution of a circumbinary disks. We do not yet allow for a dead zone in the disk.

The thermal timescale is much smaller than the viscous timescale in a circumbinary disk, and so we approximate the temperature in equilibrium.

\section{Thermodynamics}
We allow for heating of the disk due to viscous dissipation, irradiation from the binary, and dissipation of density waves due to a tidal torque.

Assuming thermal equilibrium, we have, at low optical depths, $\tau \ll 1$,

\begin{equation}
\tau \sigma T_m^4 = F_{tid} + F_{\nu} + \tau F_{irr}\,,
\end{equation}

and at high optical depths, $\tau \gg 1$,

\begin{equation}
\sigma T_m^4 = F_{tid} + \frac{3\tau}{4}  F_\nu + F_{irr}\,.
\end{equation}

Extrapolated to all optical depths,

\begin{equation} \label{eq:encon}
\sigma T_m^4 = \left(1+\frac{1}{\tau}\right) F_{tid} + \left(\frac{3\tau}{4} + \frac{1}{\tau}\right) F_{v} + F_{irr}\,,
\end{equation}

where $T_m$ the disk midplane temperature. The optical depth is defined as $\tau =\frac{1}{2} \kappa \Sigma$, where $\kappa$ is the opacity. We will assume a power-law for the opacity,

\begin{equation}
\kappa = \kappa_o T^\beta
\end{equation} 

The tidal heating term is \textbf{($F_{\nu}$ is derived per disk face. Is $F_{tid}$ also per disk face or should we divide by 2?})

\begin{equation}
F_{tid} = \left(\Omega_b - \Omega\right) \Lambda \Sigma\,,
\end{equation}

where $\Omega_b = \sqrt{G M/a^3}$ is the binary angular velocity, with $a$ the binary semimajor axis, $\Omega = \sqrt{G M/r^{3}}$ is the Keplerian velocity, $M$ is the total mass of the binary, and $\Lambda$ is the torque due to the binary. At radius $r > h(r) + a$,

\begin{equation}
\Lambda = f \frac{q^2 GM}{a} \left(\frac{a}{R-a}\right)^4\,,
\end{equation}

where $q$ is the binary mass fraction and $f$ is a correction factor that fixes the inner disk truncation radius. Note that Martin 2013 fix $f =1$, which places their disk inner edge further out. 

The disk height $h(r)$ is given by

\begin{equation}
\frac{h}{r} = \left(\frac{k\,T\,r}{G M\mu}\right)^\frac{1}{2} \,,
\end{equation}

by assuming hydrostatic equilibrium for a gas-pressure dominated disk.
The heating due to irradiation from the binary is

\begin{equation}
F_{irr} = \frac{1}{2} \frac{L_c}{4\pi r^2} \zeta\,,
\end{equation}

where $\zeta$ is the incidence angle of radiation from the binary onto the disk at a height $\eta h$. Far from the binary, we can approximate $\zeta$ as

\begin{equation}
\zeta = r \frac{d}{dr} \left(\frac{\eta h}{r}\right)\,.
\end{equation}

The heating due to viscous dissipation is

\begin{equation}
F_\nu = \frac{3}{8\pi} \frac{F_J \Omega}{r^2} = \frac{9}{8} \Omega^2 \nu \Sigma\,,
\end{equation}

where the second equality comes from the definition of the angular momentum flux $F_J = 3 \pi \nu \Sigma l$, with $l = \Omega r^2$, the specific angular momentum. (\textit{adjust for dead zone}) The viscosity is expressed as

\begin{equation}
\nu = \alpha \frac{c_s^2}{\Omega}
\end{equation}

where the sound speed is $c_s = \sqrt{\frac{P}{\rho}}$ and the viscosity is for a gas-pressure dominated disk 
 
\begin{equation}
v = 	\alpha \frac{k_B T}{\mu \Omega}\,.
\end{equation}

We can simplify the energy balance equation, Eq.\,\ref{eq:encon}, by solving for the temperature in the limit of various heating processes and then interpolating our results as

\begin{equation}
T_m^4 = \left(T_{tid,thin}^4 + T_{tid, thick}^4\right)  + \left(T_{\nu,thin}^4 + T_{\nu,thick}^4\right) + T_{irr}^4\,,
\end{equation}

where, for example, we find $T_{irr}$ by equating $F_{irr} = \sigma T_{irr}^4$. (\textbf{We will check for error for this approximation by solving Eq. 3 exactly for a few timesteps. Also, we can probably perform stability tests on the solution})

In the limit of small optical depth, $\tau \ll 1$, and heating dominated by viscous dissipation, we have

\begin{equation}
T _{\nu,thin} = \left(\frac{9}{4} \frac{\alpha k_b}{\sigma \mu \kappa_o}\Omega \right)^\frac{1}{3+\beta}\,.
\end{equation}

In the limit of large optical depth, $\tau \gg 1$, and heating dominated by viscous dissipation, we have

\begin{equation}
T_{\nu,thick}= \left(\frac{27}{64} \frac{\kappa_o \alpha k_b}{\sigma \mu} \Omega\,\Sigma^2\right)^\frac{1}{3-\beta}\,.
\end{equation}

In the limit of small optical depth and heating dominated by tidal dissipation, we have

\begin{equation}
T_{tid,thin} = \left(\frac{2}{\sigma \kappa_o}(\Omega_b - \Omega)\, \Lambda \right)^\frac{1}{4+\beta}\,.
\end{equation}

In the limit of large optical depth and heating dominated by tidal dissipation, we have

\begin{equation}
T_{tid,thick} = \left(\frac{1}{\sigma}(\Omega_b - \Omega)\,\Lambda\,\Sigma\right)^\frac{1}{4}\,.
\end{equation}

For heating dominated by irradiation, we have

\begin{equation}
T_{irr} = \left(\left(\frac{\eta}{7}\frac{L_c}{4\pi\sigma}\right)^2 \frac{k_b}{\mu G M}\right)^\frac{1}{7} r^{-\frac{3}{7}}\,.
\end{equation}

\section{Disk Evolution}

We can now evaluate the disk evolution equation by expressing viscosity as a function of surface density $\Sigma$ and radius $r$.

The disk evolution equation is

\begin{equation} \label{eq:diskev}
\frac{\partial \Sigma}{\partial t} = \frac{1}{r} \frac{\partial}{\partial r} \left[3 r^{1/2} \frac{\partial}{\partial r} \left(r^{1/2} \nu \Sigma\right) - \frac{2 \Lambda \Sigma r^{3/2}}{(G M)^{1/2}}\right]\,.
\end{equation}

(\textit{the writeup is missing the factor of 2 in front of the external torque term})


\section{Conservation Equations}

The equation for mass conservation is

\begin{equation}
M = \int 2\pi r \Sigma\, dr\,,
\end{equation}

and for momentum conservation,

\begin{equation}
\textit{L} = \int2\pi r \Sigma\, r^2 \Omega\, dr\,.
\end{equation}

We express the evolution equation in a conservative form, in terms of $\Sigma$ rather than $F_J$. This will facilitate verifying conservation every time step of the numerical procedure.

\section{Initial and Boundary Conditions}

We will set $\Sigma = 0$ for our inner boundary condition. Our outer boundary condition is irrelevant because our solution will never reach it, so we can likewise set $\Sigma = 0$ at our outer boundary condition. 

We can derive a more general inner boundary condition with the hypothesis $\partial \dot{M}/ \partial l = 0$, or equivalently,

\begin{equation}
\frac{\partial}{\partial l} \left(3 r^{1/2} \frac{\partial}{\partial r} \left(r^{1/2} \nu \Sigma\right) - \frac{2 \Lambda \Sigma r^{3/2}}{(G M)^{1/2}}\right)= 0\,.
\end{equation}

The binary torque $\Lambda$ vanishes to 0 at the inner edge. If we further assume that the accretion rate $\dot{M}$ is 0, we get that $\Sigma$ must be a constant, which we can fix to 0 and return our inner boundary condition above.

The initial condition is yet undetermined.

\section{Numerical Method}

We will use FiPy to solve Eq.\,\ref{eq:diskev}. The package supports partial differential equations with a nonlinear diffusion term and an advection term.

\end{document}